---
title: "Mock — lung exceedance plot"
output:
  html_document:
    toc: false
    df_print: paged
params:
  site_code: ""
---

Experimental plot: simple lung outlines filled like the annual exceedance bar charts.

- **Left lung** — NO~2~ exceedance (% of days + day count), colour `#7e0052`
- **Right lung** — PM~2.5~ exceedance (% of days + day count), black fill
- One lung pair **per calendar year** (same data as the Summary-of-data-to-date bar charts)

This file is separate from `QuarterlyAQtrends_git.Rmd` — tweak here before moving anything into the main report.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

if (!file.exists("R/bl_api.R")) {
  stop("Knit from the project root (folder containing R/bl_api.R).")
}

source("R/bl_api.R")
source("R/bl_report_stats.R")
source("R/bl_lung_exceedance_plot.R")
bl_load_env()

suppressPackageStartupMessages({
  library(dplyr)
  library(ggplot2)
})

site_code <- params$site_code
if (is.null(site_code) || !nzchar(site_code)) {
  site_code <- bl_resolve_report_site_codes()[1]
}
bl_use_site_code(site_code)

WHO_NO2_DAILY <- 25
WHO_PM25_DAILY <- 15

report_data <- bl_load_report_data(site_code = site_code)
report_dates <- bl_env_report_dates(site_code = site_code)

no2_total_stats <- bl_to_date_pollutant_summary(
  report_data$no2, "no2", WHO_NO2_DAILY,
  report_dates$quarter_end, report_dates$report_year, report_dates$report_quarter
)
pm25_total_stats <- bl_to_date_pollutant_summary(
  report_data$pm25, "pm25", WHO_PM25_DAILY,
  report_dates$quarter_end, report_dates$report_year, report_dates$report_quarter
)
```

## Site: `r site_code`

```{r lung-plot, fig.width=11, fig.height=4.5, fig.align='center'}
lung_plot <- bl_plot_lung_exceedance(
  no2_total_stats,
  pm25_total_stats,
  site_code = site_code
)

if (is.null(lung_plot)) {
  knitr::knit_exit("No annual exceedance data available for lung plot.")
}

print(lung_plot)
```

## Underlying data

```{r data-table}
bl_lung_exceedance_year_df(no2_total_stats, pm25_total_stats)
```

## Compare with bar charts (simple columns — full bars are in the main report)

```{r bar-no2, fig.width=7, fig.height=3, fig.align='center'}
plot_df <- no2_total_stats$exceed_days_by_year
ggplot(plot_df, aes(x = factor(year), y = exceed_pct)) +
  geom_col(fill = "#7e0052", width = 0.6) +
  geom_text(aes(label = paste0(exceed_pct, "%")), vjust = -0.3, size = 3.5) +
  labs(title = "NO2 exceedance % (reference)", x = NULL, y = "% days") +
  theme_minimal()
```

```{r bar-pm25, fig.width=7, fig.height=3, fig.align='center'}
plot_df <- pm25_total_stats$exceed_days_by_year
ggplot(plot_df, aes(x = factor(year), y = exceed_pct)) +
  geom_col(fill = "black", width = 0.6) +
  geom_text(aes(label = paste0(exceed_pct, "%")), vjust = -0.3, size = 3.5) +
  labs(title = "PM2.5 exceedance % (reference)", x = NULL, y = "% days") +
  theme_minimal()
```
