---
title: "Breathe London Community Programme"
author: |
  | Dr. Kayla Schulte, Environmental Research Group, Imperial College London
  | k.schulte21@imperial.ac.uk
output: pdf_document
date: "`r format(Sys.Date(), '%d %b %Y')`" # Today's date
colorlinks: true
header-includes:
  - \usepackage{helvet}
  - \renewcommand{\familydefault}{\sfdefault}
  - \usepackage{xcolor}
  - \definecolor{blurl}{HTML}{409894}
  - \definecolor{blgrey}{HTML}{888888}
  - \AtBeginDocument{\hypersetup{urlcolor=blurl, linkcolor=blurl}}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

if (!file.exists("R/bl_api.R")) {
  stop("Knit from the project root (folder containing R/bl_api.R).")
}
source("R/bl_api.R")
bl_load_env()

packages <- c(
  "dplyr", "ggplot2", "tidyverse", "cowplot", "tidyr",
  "jsonlite", "stringr", "lubridate"
)
invisible(lapply(packages, require, character.only = TRUE))

# BL_DATA_SOURCE=csv (default) or api — see .env.example
report_data <- bl_load_report_data()
NO2_DATA_FRAME <- report_data$no2
PM25_DATA_FRAME <- report_data$pm25

report_dates <- bl_env_report_dates()
quarter_start <- report_dates$quarter_start
quarter_end <- report_dates$quarter_end
report_year <- report_dates$report_year
report_quarter <- report_dates$report_quarter

quarter_headings <- c(
  "Quarter 1 (1 January to 31 March)",
  "Quarter 2 (1 April to 30 June)",
  "Quarter 3 (1 July to 30 September)",
  "Quarter 4 (1 October to 31 December)"
)
quarter_heading <- paste(quarter_headings[report_quarter], "in", report_year)
quarter_label <- paste0("Q", report_quarter)
plot_legend_labels <- c("All data to date", quarter_label)

quarter_trend_texts <- c(
  "(Q1) Jan-March: In this quarter we may see higher NO$_2$ levels if we have some cold winter mornings that trap emissions near the ground. Towards the end of March we may see some particulate episodes as the farming fertiliser season starts and we have some still high-pressure conditions.",
  "(Q2) April-June: We may continue to see some particulate episodes in April as the farming fertiliser season continues and we have high pressure systems creating stagnant conditions. May and June are generally low pollution months.",
  "(Q3) July-Sept: The summer months are when we are likely to see ozone episodes. Ozone is not emitted but rather formed in the atmosphere from other pollutants in the heat of the sun. Generally, ozone (O$_3$) builds through the day peaking between 16:00-20:00.",
  "(Q4) Oct-Dec: Autumn is typically a season of lower pollution, with the wind and rain helping to keep pollution low. As we get into December though we do sometimes see winter smogs on cold, clear, crisp mornings where pollution from traffic and heating is trapped near the ground."
)

bl_format_quarter_line <- function(q, text) {
  if (q == report_quarter) {
    sprintf("\\textcolor{blurl}{\\textbf{%s}}", text)
  } else {
    sprintf("\\textcolor{blgrey}{%s}", text)
  }
}

```

\centering
## Quarterly Air Quality Trends Report


# Mums for Lungs Redbridge

\raggedright

## `r quarter_heading`

This document presents hourly average pollution concentration measured at your Breathe London Community Programme node over the past 3 months. 

## General Air Pollution Trends in London 

```{r quarter-trends, echo=FALSE, results='asis'}
for (q in seq_along(quarter_trend_texts)) {
  cat(bl_format_quarter_line(q, quarter_trend_texts[q]), "\n\n")
}
```


```{r, echo=FALSE}
#Clean raw data for graphing - Filter by weekdays vs weekends for both pollutants
#IF USING EXPORTED CSV FROM BREATHELONDON.ORG - RENAME COLUMNS
#NO2_DATA_FRAME <- NO2_DATA_FRAME %>% 
#  rename(
#    date = Category,
#    no2 = 'Nitrogen dioxide'
#    )

#PM25_DATA_FRAME <- PM25_DATA_FRAME %>% 
#  rename(
#    date = Category,
#    pm25 = 'PM<sub>2.5</sub> particulates'
#    )

#NO2

NO2_raw <- NO2_DATA_FRAME
NO2_raw$date <- gsub("\\..*","",NO2_raw$date)

NO2_raw$weekday <- lubridate::wday(lubridate::ymd_hms(NO2_raw$date))

NO2_weekdays <- NO2_raw %>% filter(
  weekday == 2 | weekday == 3 | weekday == 4 | weekday == 5 | weekday == 6
)

NO2_weekends <- NO2_raw %>% filter(
  weekday == 7 | weekday == 1
)

#PM2.5 

PM2.5_raw <- PM25_DATA_FRAME
PM2.5_raw$date <- gsub("\\..*","",PM2.5_raw$date)

PM2.5_raw$weekday <- lubridate::wday(lubridate::ymd_hms(PM2.5_raw$date))

PM2.5_weekdays <- PM2.5_raw %>% filter(
  weekday == 2 | weekday == 3 | weekday == 4 | weekday == 5 | weekday == 6
)

PM2.5_weekends <- PM2.5_raw %>% filter(
  weekday == 7 | weekday == 1
)

#Calculate hourly means by weekday
#NO2

NO2_weekdays <- NO2_weekdays %>%
  dplyr::mutate(time = sub("^\\S+\\s+", "", date))

NO2_weekdays <- separate(NO2_weekdays, date, into = c("Date2", "time2"), sep = "^\\S*\\K\\s+")
NO2_weekdays <- subset(NO2_weekdays, select = -c(time2))
NO2_weekdays$Date2 <- as.Date(NO2_weekdays$Date2)

NO2_weekdays <- within(NO2_weekdays,{
  posb <- as.POSIXlt(time,format="%H:%M:%S")
  hours <- posb$hour
  mins <- posb$min
  time <- format(posb, "%H:%M")
  posb <- NULL  # cleanup
})

hours2 <- mapply(function(x, y) paste0(rep(x, y), collapse = ""), 0, 4 - nchar(NO2_weekdays$hours))
hours <- paste0(hours2, NO2_weekdays$hours)
hours <- format(strptime(hours, format="%M%H"), format = "%H:%M")
NO2_weekdays$hours <- hours

NO2_weekdays_Q3 <- NO2_weekdays %>% filter(between(Date2, report_dates$q3_start, report_dates$q3_end)) 
NO2_weekdays_quarter <- NO2_weekdays %>% filter(between(Date2, quarter_start, quarter_end))
NO2_weekdays_all <- NO2_weekdays

NO2_weekdays_means_Q3 <- NO2_weekdays_Q3 %>% group_by(hours) %>%                        
  summarise_at(vars(no2),
               list(mean = mean))

NO2_weekdays_means_quarter <- NO2_weekdays_quarter %>% group_by(hours) %>%                        
  summarise_at(vars(no2),
               list(mean = mean)) 

NO2_weekdays_means_all <- NO2_weekdays_all %>% group_by(hours) %>%                        
  summarise_at(vars(no2),
               list(mean = mean)) 

#PM2.5

PM2.5_weekdays <- PM2.5_weekdays %>%
  dplyr::mutate(time = sub("^\\S+\\s+", "", date))

PM2.5_weekdays <- separate(PM2.5_weekdays, date, into = c("Date2", "time2"), sep = "^\\S*\\K\\s+")
PM2.5_weekdays <- subset(PM2.5_weekdays, select = -c(time2))
PM2.5_weekdays$Date2 <- as.Date(PM2.5_weekdays$Date2)

PM2.5_weekdays <- within(PM2.5_weekdays,{
  posb <- as.POSIXlt(time,format="%H:%M:%S")
  hours <- posb$hour
  mins <- posb$min
  time <- format(posb, "%H:%M")
  posb <- NULL  # cleanup
})

hours2 <- mapply(function(x, y) paste0(rep(x, y), collapse = ""), 0, 4 - nchar(PM2.5_weekdays$hours))
hours <- paste0(hours2, PM2.5_weekdays$hours)
hours <- format(strptime(hours, format="%M%H"), format = "%H:%M")
PM2.5_weekdays$hours <- hours

PM2.5_weekdays_Q3 <- PM2.5_weekdays %>% filter(between(Date2, report_dates$q3_start, report_dates$q3_end)) 
PM2.5_weekdays_quarter <- PM2.5_weekdays %>% filter(between(Date2, quarter_start, quarter_end))
PM2.5_weekdays_all <- PM2.5_weekdays

PM2.5_weekdays_means_Q3 <- PM2.5_weekdays_Q3 %>% group_by(hours) %>%                        
  summarise_at(vars(pm25),
               list(mean = mean))

PM2.5_weekdays_means_quarter <- PM2.5_weekdays_quarter %>% group_by(hours) %>%                        
  summarise_at(vars(pm25),
               list(mean = mean))

PM2.5_weekdays_means_all <- PM2.5_weekdays_all %>% group_by(hours) %>%                        
  summarise_at(vars(pm25),
               list(mean = mean))



#Generate  dfs to plot pollution trends

#NO2

NO2_weekdays_means_all$name <- "All data to date"
NO2_weekdays_means_Q3$name <- "Q3"
NO2_weekdays_means_quarter$name <- quarter_label

NO2_plot_df <- rbind(NO2_weekdays_means_all, NO2_weekdays_means_quarter)

#PM2.5

PM2.5_weekdays_means_all$name <- "All data to date"
PM2.5_weekdays_means_Q3$name <- "Q3"
PM2.5_weekdays_means_quarter$name <- quarter_label

PM2.5_plot_df <- rbind(PM2.5_weekdays_means_all, PM2.5_weekdays_means_quarter)

#Calculate hourly means by weekend day
#NO2

NO2_weekends <- NO2_weekends %>%
  dplyr::mutate(time = sub("^\\S+\\s+", "", date))

NO2_weekends <- separate(NO2_weekends, date, into = c("Date2", "time2"), sep = "^\\S*\\K\\s+")
NO2_weekends <- subset(NO2_weekends, select = -c(time2))
NO2_weekends$Date2 <- as.Date(NO2_weekends$Date2)

NO2_weekends <- within(NO2_weekends,{
  posb <- as.POSIXlt(time,format="%H:%M:%S")
  hours <- posb$hour
  mins <- posb$min
  time <- format(posb, "%H:%M")
  posb <- NULL  # cleanup
})

hours2 <- mapply(function(x, y) paste0(rep(x, y), collapse = ""), 0, 4 - nchar(NO2_weekends$hours))
hours <- paste0(hours2, NO2_weekends$hours)
hours <- format(strptime(hours, format="%M%H"), format = "%H:%M")
NO2_weekends$hours <- hours

NO2_weekends_Q3 <- NO2_weekends %>% filter(between(Date2, report_dates$q3_start, report_dates$q3_end)) 
NO2_weekends_quarter <- NO2_weekends %>% filter(between(Date2, quarter_start, quarter_end))
NO2_weekends_all <- NO2_weekends

NO2_weekends_means_Q3 <- NO2_weekends_Q3 %>% group_by(hours) %>%                        
  summarise_at(vars(no2),
               list(mean = mean))

NO2_weekends_means_quarter <- NO2_weekends_quarter %>% group_by(hours) %>%                        
  summarise_at(vars(no2),
               list(mean = mean)) 

NO2_weekends_means_all <- NO2_weekends_all %>% group_by(hours) %>%                        
  summarise_at(vars(no2),
               list(mean = mean)) 

#PM2.5

PM2.5_weekends <- PM2.5_weekends %>%
  dplyr::mutate(time = sub("^\\S+\\s+", "", date))

PM2.5_weekends <- separate(PM2.5_weekends, date, into = c("Date2", "time2"), sep = "^\\S*\\K\\s+")
PM2.5_weekends <- subset(PM2.5_weekends, select = -c(time2))
PM2.5_weekends$Date2 <- as.Date(PM2.5_weekends$Date2)

PM2.5_weekends <- within(PM2.5_weekends,{
  posb <- as.POSIXlt(time,format="%H:%M:%S")
  hours <- posb$hour
  mins <- posb$min
  time <- format(posb, "%H:%M")
  posb <- NULL  # cleanup
})

hours2 <- mapply(function(x, y) paste0(rep(x, y), collapse = ""), 0, 4 - nchar(PM2.5_weekends$hours))
hours <- paste0(hours2, PM2.5_weekends$hours)
hours <- format(strptime(hours, format="%M%H"), format = "%H:%M")
PM2.5_weekends$hours <- hours

PM2.5_weekends_Q3 <- PM2.5_weekends %>% filter(between(Date2, report_dates$q3_start, report_dates$q3_end)) 
PM2.5_weekends_quarter <- PM2.5_weekends %>% filter(between(Date2, quarter_start, quarter_end))
PM2.5_weekends_all <- PM2.5_weekends

PM2.5_weekends_means_Q3 <- PM2.5_weekends_Q3 %>% group_by(hours) %>%                        
  summarise_at(vars(pm25),
               list(mean = mean))

PM2.5_weekends_means_quarter <- PM2.5_weekends_quarter %>% group_by(hours) %>%                        
  summarise_at(vars(pm25),
               list(mean = mean))

PM2.5_weekends_means_all <- PM2.5_weekends_all %>% group_by(hours) %>%                        
  summarise_at(vars(pm25),
               list(mean = mean))



#Generate  dfs to plot pollution trends

#NO2

NO2_weekends_means_all$name <- "All data to date"
NO2_weekends_means_Q3$name <- "Q3"
NO2_weekends_means_quarter$name <- quarter_label

NO2_plot_weekends_df <- rbind(NO2_weekends_means_all, NO2_weekends_means_quarter)

#PM2.5

PM2.5_weekends_means_all$name <- "All data to date"
PM2.5_weekends_means_Q3$name <- "Q3"
PM2.5_weekends_means_quarter$name <- quarter_label

PM2.5_plot_weekends_df <- rbind(PM2.5_weekends_means_all, PM2.5_weekends_means_quarter)

```

# NO~2~ 

## Average NO~2~ each hour at Breathe London node

```{r, out.width='\\textwidth', fig.width = 10, fig.height = 6, fig.align='center', fig.show='hold'}
NO2_plot <- ggplot(NO2_plot_df)+
  geom_bar(aes(factor(hours),mean,fill=name),stat="identity",position="identity",alpha=0.9)+
  theme_minimal() + theme(axis.text.x = element_text(angle = 45, vjust = 0.8, hjust=1)) + 
  xlab("Hour (Weekday)") + ylab("Mean NO"[2]~"in ug/u3") + ggtitle("WEEKDAYS") +
  geom_hline(yintercept = 40, linetype = "dashed", color = "black") +
  annotate("text", x = 0, y = 40, label = "UK NO2 annual mean objective (40 ug/m3)", vjust = 1.5, hjust = -0.1, color = "black", size = 3.5) +
  geom_hline(yintercept = 25, linetype = "dashed", color = "black") +
  annotate("text", x = 0, y = 25, label = "WHO NO2 daily mean objective (25 ug/m3)", vjust = 1.5, hjust = -0.1, color = "black", size = 3.5) +
  geom_hline(yintercept = 5, linetype = "dashed", color = "black") +
  annotate("text", x = 0, y = 5, label = "WHO NO2 annual mean objective (10 ug/m3)", vjust = 1.5, hjust = -0.1, color = "black", size = 3.5) + 
  scale_fill_manual(values=c("#038476", "#A7D5AF"), name = "Data range", labels = plot_legend_labels) +
  theme(legend.position="none", plot.title = element_text(hjust = 0.5))

NO2_plot_weekend <- ggplot(NO2_plot_weekends_df)+
  geom_bar(aes(factor(hours),mean,fill=name),stat="identity",position="identity",alpha=0.9)+
  theme_minimal() + theme(axis.text.x = element_text(angle = 45, vjust = 0.8, hjust=1)) + 
  xlab("Hour (Weekend)") + ggtitle("WEEKENDS") +
  geom_hline(yintercept = 40, linetype = "dashed", color = "black") +
  annotate("text", x = 0, y = 40, label = "UK NO2 annual mean objective (40 ug/m3)", vjust = 1.5, hjust = -0.1, color = "black", size = 3.5) +
  geom_hline(yintercept = 25, linetype = "dashed", color = "black") +
  annotate("text", x = 0, y = 25, label = "WHO NO2 daily mean objective (25 ug/m3)", vjust = 1.5, hjust = -0.1, color = "black", size = 3.5) +
  geom_hline(yintercept = 5, linetype = "dashed", color = "black") +
  annotate("text", x = 0, y = 5, label = "WHO NO2 annual mean objective (10 ug/m3)", vjust = 1.5, hjust = -0.1, color = "black", size = 3.5) + 
  scale_fill_manual(values=c("#038476", "#A7D5AF"), name = "Data range", labels = plot_legend_labels) +
   theme(plot.title = element_text(hjust = 0.5))

plot_grid(NO2_plot, NO2_plot_weekend, align = "h", ncol = 2, rel_widths = c(1/4, 1/3))
```

# PM~2.5~

## Average PM~2.5~ each hour at Breathe London node

```{r, out.width='\\textwidth', fig.width = 10, fig.height = 6, fig.align='center', fig.show='hold'}
PM2.5_plot_weekdays <- ggplot(PM2.5_plot_df)+
  geom_bar(aes(factor(hours),mean,fill=name),stat="identity",position="identity",alpha=0.9)+
  theme_minimal() + theme(axis.text.x = element_text(angle = 45, vjust = 0.8, hjust=1)) + 
  xlab("Hour") + ylab("Mean PM2.5 in ug/u3") + ggtitle("WEEKDAYS") +
  geom_hline(yintercept = 25, linetype = "dashed", color = "black") +
  annotate("text", x = 0, y = 25, label = "UK PM2.5 annual mean objective (25 ug/m3)", vjust = 1.5, hjust = -0.1, color = "black", size = 3.5) +
  geom_hline(yintercept = 15, linetype = "dashed", color = "black") +
  annotate("text", x = 0, y = 15, label = "WHO PM2.5 daily mean objective (15 ug/m3)", vjust = 1.5, hjust = -0.1, color = "black", size = 3.5) +
  geom_hline(yintercept = 5, linetype = "dashed", color = "black") +
  annotate("text", x = 0, y = 5, label = "WHO PM2.5 annual mean objective (5 ug/m3)", vjust = 1.5, hjust = -0.1, color = "black", size = 3.5) +
  scale_fill_manual(values=c("#038476", "#A7D5AF"), name = "Data range", labels = plot_legend_labels) +
  theme(legend.position="none", plot.title = element_text(hjust = 0.5))

PM2.5_plot_weekends <- ggplot(PM2.5_plot_weekends_df)+
  geom_bar(aes(factor(hours),mean,fill=name),stat="identity",position="identity",alpha=0.9)+
  theme_minimal() + theme(axis.text.x = element_text(angle = 45, vjust = 0.8, hjust=1)) + 
  xlab("Hour") + ggtitle("WEEKENDS") +
  geom_hline(yintercept = 25, linetype = "dashed", color = "black") +
  annotate("text", x = 0, y = 25, label = "UK PM2.5 annual mean objective (25 ug/m3)", vjust = 1.5, hjust = -0.1, color = "black", size = 3.5) +
  geom_hline(yintercept = 15, linetype = "dashed", color = "black") +
  annotate("text", x = 0, y = 15, label = "WHO PM2.5 daily mean objective (15 ug/m3)", vjust = 1.5, hjust = -0.1, color = "black", size = 3.5) +
  geom_hline(yintercept = 5, linetype = "dashed", color = "black") +
  annotate("text", x = 0, y = 5, label = "WHO PM2.5 annual mean objective (5 ug/m3)", vjust = 1.5, hjust = -0.1, color = "black", size = 3.5) +
  scale_fill_manual(values=c("#038476", "#A7D5AF"), name = "Data range", labels = plot_legend_labels)+
   theme(plot.title = element_text(hjust = 0.5))

plot_grid(PM2.5_plot_weekdays, PM2.5_plot_weekends, align = "h", ncol = 2, rel_widths = c(1/4, 1/3)) 

```

To download the data that was used to generate this report, visit <https://www.breathelondon.org/> and search for your node in the Node Search hyperlink just below the node map. 


