---
title: "Breathe London Community Programme"
knit: (function(inputFile, ...) { source("scripts/knit_quarterly_report.R"); knit_quarterly_report(inputFile, ...) })
params:
  site_code: ""
  # When TRUE, knit only the NO2 or PM2.5 hourly plot (fast layout iteration)
  preview_no2_hourly: false
  preview_pm25_hourly: false
output: pdf_document
colorlinks: true
header-includes:
  - \usepackage[T1]{fontenc}
  - \usepackage{helvet}
  - \usepackage{sansmath}
  - \renewcommand{\familydefault}{\sfdefault}
  - \AtBeginDocument{\sansmath}
  - \usepackage{xcolor}
  - \usepackage{graphicx}
  - \usepackage{float}
  - \usepackage{tikz}
  - \usepackage{anyfontsize}
  - \usepackage{framed}
  - \input{R/bl_summary_box.tex}
  - \definecolor{blurl}{HTML}{0000CD}
  - \definecolor{blgrey}{HTML}{888888}
  - \definecolor{blorg}{HTML}{0000CD}
  - \definecolor{blno2totalborder}{HTML}{7e0052}
  - \definecolor{blno2quarterborder}{HTML}{eea7e6}
  - \definecolor{blpm25totalborder}{HTML}{000000}
  - \definecolor{blpm25quarterborder}{HTML}{B0B0C0}
  - \definecolor{blnearest}{HTML}{069494}
  - \setlength{\fboxsep}{10pt}
  - \setlength{\fboxrule}{3pt}
  - \AtBeginDocument{\hypersetup{urlcolor=blurl, linkcolor=blurl}}
  - \makeatletter
  - \newcommand{\blpageonefooter}{}
  - \input{R/bl_title_banner.tex}
  - \def\ps@blfirstpage{\def\@oddfoot{\blpageonefooter}\let\@evenfoot\@oddfoot\def\@oddhead{}\let\@evenhead\@oddhead}
  - \makeatother
  - \renewcommand{\maketitle}{}
  - \raggedbottom
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

if (!file.exists("R/bl_api.R")) {
  stop("Knit from the project root (folder containing R/bl_api.R).")
}
source("R/bl_api.R")
preview_no2 <- isTRUE(params$preview_no2_hourly)
preview_pm25 <- isTRUE(params$preview_pm25_hourly)
preview_hourly <- preview_no2 || preview_pm25
if (!preview_hourly) {
  source("R/bl_marylebone.R")
  source("R/bl_report_map.R")
  source("R/bl_lung_exceedance_plot.R")
}
bl_load_env()

site_code <- params$site_code
if (is.null(site_code) || !nzchar(site_code)) {
  site_code <- bl_resolve_report_site_codes()[1]
}
bl_use_site_code(site_code)

packages <- c(
  "dplyr", "ggplot2", "cowplot", "tidyr", "lubridate"
)
invisible(lapply(packages, require, character.only = TRUE))

# BL_DATA_SOURCE=csv (default) or api — see .env.example
report_data <- bl_load_report_data(site_code = site_code)
NO2_DATA_FRAME <- report_data$no2
PM25_DATA_FRAME <- report_data$pm25

report_dates <- bl_env_report_dates(site_code = site_code)
organisation_name <- bl_site_organisation_name(site = site_code)
quarter_start <- report_dates$quarter_start
quarter_end <- report_dates$quarter_end
measurement_start <- report_dates$fetch_start
report_year <- report_dates$report_year
report_quarter <- report_dates$report_quarter

quarter_label <- paste0("Q", report_quarter)
summary_to_date_heading <- sprintf(
  "Summary of data to date (%s to present)",
  sprintf(
    "%d %s %s",
    lubridate::day(measurement_start),
    format(measurement_start, "%B"),
    format(measurement_start, "%Y")
  )
)
report_title_line1 <- sprintf("Quarter %d", report_quarter)
report_title_line2 <- "Air Quality Trends Report"
report_title_line3 <- sprintf(
  "%s to %s %s",
  sprintf("%d %s", lubridate::day(quarter_start), format(quarter_start, "%B")),
  sprintf("%d %s", lubridate::day(quarter_end), format(quarter_end, "%B")),
  report_year
)
all_data_label <- "All data to date"

quarter_trend_texts <- c(
  "(Q1) Jan-March: In this quarter we may see higher NO\\textsubscript{2} levels if we have some cold winter mornings that trap emissions near the ground. Towards the end of March we may see some particulate episodes as the farming fertiliser season starts and we have some still high-pressure conditions.",
  "(Q2) April-June: We may continue to see some particulate episodes in April as the farming fertiliser season continues and we have high pressure systems creating stagnant conditions. May and June are generally low pollution months.",
  "(Q3) July-Sept: The summer months are when we are likely to see ozone episodes. Ozone is not emitted but rather formed in the atmosphere from other pollutants in the heat of the sun. Generally, ozone (O\\textsubscript{3}) builds through the day peaking between 4-8 PM.",
  "(Q4) Oct-Dec: Autumn is typically a season of lower pollution, with the wind and rain helping to keep pollution low. As we get into December though we do sometimes see winter smogs on cold, clear, crisp mornings where pollution from traffic and heating is trapped near the ground."
)

bl_format_quarter_line <- function(q, text) {
  if (q == report_quarter) {
    sprintf("\\textcolor{blorg}{\\textbf{%s}}", text)
  } else {
    sprintf("\\textcolor{blgrey}{%s}", text)
  }
}

WHO_NO2_DAILY <- 25
WHO_PM25_DAILY <- 15
BL_SPIKE_COOCCURRENCE_QUANTILE <- 0.75
BL_REPORT_ROUND_DIGITS <- 0L

bl_typical_peak_methodology_text <- paste(
  "Typical peak time ranges are based on the day(s) with the highest peak hourly average (within 2\\% of the maximum). The reported range spans up to five contiguous hours centred on the peak hour and includes neighbouring hours with average values at or above 92\\% of the peak."
)
bl_typical_peak_footnote_mark <- "\\textsuperscript{1}"

bl_append_typical_peak_footnote <- function(text) {
  sprintf("%s%s", text, bl_typical_peak_footnote_mark)
}

bl_pollutant_summary <- function(df, value_col, threshold, start_date = NULL, end_date = NULL) {
  period_df <- df %>%
    mutate(day = as.Date(sub(" .*", "", gsub("\\..*", "", date))))

  if (!is.null(start_date)) {
    period_df <- period_df %>% filter(day >= start_date)
  }
  if (!is.null(end_date)) {
    period_df <- period_df %>% filter(day <= end_date)
  }

  values <- period_df[[value_col]]
  values <- values[!is.na(values)]

  daily <- period_df %>%
    group_by(day) %>%
    summarise(
      daily_mean = round(mean(.data[[value_col]], na.rm = TRUE), BL_REPORT_ROUND_DIGITS),
      .groups = "drop"
    )

  if (!length(values)) {
    return(list(
      exceed_days = 0L,
      exceed_days_by_year = data.frame(
        year = integer(),
        exceed_days = integer(),
        total_days = integer(),
        exceed_pct = numeric()
      ),
      mean = NA_real_,
      min = NA_real_,
      max = NA_real_
    ))
  }

  exceed_days_by_year <- daily %>%
    mutate(year = lubridate::year(day)) %>%
    group_by(year) %>%
    summarise(
      exceed_days = sum(daily_mean > threshold, na.rm = TRUE),
      total_days = sum(!is.na(daily_mean)),
      .groups = "drop"
    ) %>%
    arrange(year)

  list(
    exceed_days = sum(daily$daily_mean > threshold, na.rm = TRUE),
    exceed_days_by_year = exceed_days_by_year,
    mean = round(mean(values), BL_REPORT_ROUND_DIGITS),
    min = min(values),
    max = max(values)
  )
}

# Each year's bracketed % is exceedance days as a share of all days with data
# in that calendar year (e.g. 122 of 279 days in 2024 → 43.7%).
bl_add_annual_exceedance_year_pct <- function(by_year) {
  if (!nrow(by_year)) {
    return(by_year)
  }
  by_year$exceed_pct <- ifelse(
    by_year$total_days > 0L,
    round(100 * by_year$exceed_days / by_year$total_days, BL_REPORT_ROUND_DIGITS),
    0
  )
  by_year
}

# Summary-of-data-to-date stats for a quarterly report:
# - Cap all values at quarter_end (ignore rows after the report quarter).
# - Annual exceedance breakdown lists complete prior calendar years only.
# - Report year appears in the breakdown for Q2–Q4 (partial year through quarter_end).
# - Report year is omitted for Q1 (that quarter has its own summary box).
bl_to_date_pollutant_summary <- function(
    df,
    value_col,
    threshold,
    quarter_end,
    report_year,
    report_quarter
) {
  stats <- bl_pollutant_summary(df, value_col, threshold, end_date = quarter_end)
  by_year <- stats$exceed_days_by_year
  if (!is.null(by_year) && nrow(by_year) && report_quarter == 1L) {
    by_year <- by_year %>% filter(year != as.integer(report_year))
  }
  stats$exceed_days_by_year <- bl_add_annual_exceedance_year_pct(by_year)
  stats
}

bl_format_exceedance_days <- function(n) {
  if (n == 0L) {
    return("no days")
  }
  word <- if (n == 1L) "day" else "days"
  sprintf("%d %s", n, word)
}

# Format rounded numbers for display; strip trailing zeros (10.0 → 10).
bl_format_decimal <- function(x, digits = BL_REPORT_ROUND_DIGITS) {
  if (length(x) != 1L) {
    return(vapply(x, bl_format_decimal, character(1), digits = digits))
  }
  if (is.na(x)) {
    return(NA_character_)
  }
  s <- format(round(x, digits), nsmall = digits, trim = TRUE, scientific = FALSE)
  s <- sub("(\\.[0-9]*?)0+$", "\\1", s)
  sub("\\.$", "", s)
}

# All pollutant concentrations (µg/m³) in the report use whole numbers.
bl_format_concentration <- function(x) {
  bl_format_decimal(x, digits = BL_REPORT_ROUND_DIGITS)
}

# Distances of 1 km or more read in km to one decimal place; anything closer
# reads in whole metres.
bl_format_distance <- function(dist_km) {
  if (!length(dist_km) || is.na(dist_km)) {
    return(NA_character_)
  }
  if (dist_km >= 1) {
    return(sprintf("%.1f km", dist_km))
  }
  sprintf("%d m", as.integer(round(dist_km * 1000)))
}

# Nearest-sensor note shown between hourly plots and the quarter summary box.
bl_format_nearest_sensor_note <- function(nearest_result) {
  if (is.null(nearest_result)) {
    return(NULL)
  }
  nearest <- nearest_result$nearest
  target <- nearest_result$target
  site_name <- trimws(as.character(nearest$SiteName[1]))
  distance <- bl_format_distance(as.numeric(nearest$dist_km[1]))
  direction <- bl_compass_direction_8(
    bl_earth_bearing_deg(
      as.numeric(target$Latitude[1]), as.numeric(target$Longitude[1]),
      as.numeric(nearest$Latitude[1]), as.numeric(nearest$Longitude[1])
    )
  )
  if (!nzchar(site_name) || is.na(distance) || is.na(direction)) {
    return(NULL)
  }
  sprintf(
    paste0(
      "The \\textcolor{blnearest}{next nearest node} to your community node is the %s node that is %s away towards the %s. ",
      "Please note that \\textcolor{blnearest}{nearest nodes} will have their own pollution sources ",
      "and weather conditions that will affect their measurements."
    ),
    site_name, distance, direction
  )
}

bl_emit_nearest_sensor_note <- function(nearest_sensor_note) {
  if (is.null(nearest_sensor_note)) {
    return(invisible(NULL))
  }
  cat("\\vspace{0.35\\baselineskip}\n")
  cat("\\begingroup\\fontsize{9}{11}\\selectfont\n")
  cat(nearest_sensor_note, "\\par\n", sep = "")
  cat("\\endgroup\n")
  cat("\\vspace{0.65\\baselineskip}\n\n")
}

bl_format_latex_pct <- function(pct) {
  # LaTeX treats % as a comment; emit \% in the PDF output.
  paste0(bl_format_decimal(pct, digits = BL_REPORT_ROUND_DIGITS), "\\%")
}

bl_format_concentration_stats <- function(stats, period_label) {
  if (is.na(stats$mean)) {
    if (identical(period_label, "To date")) {
      return("No data are available.")
    }
    return("No data are available for this quarter.")
  }
  sprintf(
    "The average concentration was %s µg/m³; the minimum was %s µg/m³; and the maximum was %s µg/m³",
    bl_format_concentration(stats$mean),
    bl_format_concentration(stats$min),
    bl_format_concentration(stats$max)
  )
}

bl_annual_exceedance_year_lines <- function(stats) {
  by_year <- stats$exceed_days_by_year
  if (is.null(by_year) || !nrow(by_year)) {
    return(NULL)
  }
  sprintf(
    "%d → %s (%s)",
    by_year$year,
    vapply(by_year$exceed_days, bl_format_exceedance_days, character(1)),
    vapply(by_year$exceed_pct, bl_format_latex_pct, character(1))
  )
}

bl_plot_annual_exceedance_bars <- function(stats, fill_color) {
  by_year <- stats$exceed_days_by_year
  if (is.null(by_year) || !nrow(by_year)) {
    return(NULL)
  }

  bar_half <- 0.45
  label_size <- 12
  days_label_size <- 6
  year_label_size <- 24
  grid_breaks <- c(25, 50, 75)

  plot_df <- by_year %>%
    mutate(
      year = factor(as.character(year), levels = as.character(year)),
      exceed_pct = as.numeric(exceed_pct),
      days_label = vapply(exceed_days, bl_format_exceedance_days, character(1)),
      pct_label = paste0(exceed_pct, "%")
    )

  ggplot(plot_df, aes(x = year)) +
    geom_rect(
      aes(
        xmin = as.numeric(year) - bar_half,
        xmax = as.numeric(year) + bar_half,
        ymin = exceed_pct,
        ymax = 100
      ),
      fill = "grey90",
      colour = NA
    ) +
    geom_hline(yintercept = grid_breaks, colour = "white", linewidth = 1.1) +
    geom_rect(
      aes(
        xmin = as.numeric(year) - bar_half,
        xmax = as.numeric(year) + bar_half,
        ymin = 0,
        ymax = exceed_pct
      ),
      fill = fill_color,
      colour = NA
    ) +
    geom_text(
      aes(x = year, y = exceed_pct, label = pct_label),
      vjust = -0.12,
      colour = fill_color,
      fontface = "bold",
      size = label_size
    ) +
    geom_text(
      aes(x = year, y = 100, label = days_label),
      vjust = -0.35,
      size = days_label_size
    ) +
    annotate(
      "text",
      x = 1 - bar_half - 0.12,
      y = grid_breaks,
      label = c("25%", "50%", "75%"),
      hjust = 1,
      colour = "grey65",
      size = 7
    ) +
    scale_y_continuous(
      expand = expansion(mult = c(0, 0.12)),
      limits = c(0, 100),
      breaks = grid_breaks,
      labels = NULL
    ) +
    scale_x_discrete(expand = expansion(add = c(1.1, 1.1))) +
    coord_cartesian(clip = "off") +
    labs(x = NULL, y = NULL) +
    theme_minimal(base_size = year_label_size) +
    theme(
      legend.position = "none",
      axis.title = element_blank(),
      axis.text.x = element_text(colour = "black", size = year_label_size),
      axis.text.y = element_blank(),
      axis.ticks.y = element_blank(),
      axis.ticks.x = element_blank(),
      axis.line.x = element_blank(),
      panel.grid = element_blank(),
      panel.background = element_blank(),
      plot.margin = margin(10, 2, 5, 2)
    )
}

bl_format_who_exceedance <- function(stats, pollutant_md, threshold, period_label) {
  if (is.na(stats$mean)) {
    if (identical(period_label, "To date")) {
      return(sprintf(
        "No WHO %s exceedance data are available.",
        pollutant_md
      ))
    }
    return(sprintf(
      "No WHO %s exceedance data are available for this quarter.",
      pollutant_md
    ))
  }
  if (stats$exceed_days == 0L) {
    return(sprintf(
      "No days exceeded the WHO %s daily mean guideline of %d µg/m³.",
      pollutant_md, threshold
    ))
  }
  sprintf(
    "%s exceeded the WHO %s daily mean guideline of %d µg/m³.",
    bl_format_exceedance_days(stats$exceed_days), pollutant_md, threshold
  )
}

bl_emit_annual_exceedance_sub_bullets <- function(stats, period_label) {
  if (!identical(period_label, "To date")) {
    return(invisible(NULL))
  }
  year_lines <- bl_annual_exceedance_year_lines(stats)
  if (is.null(year_lines)) {
    return(invisible(NULL))
  }
  cat("\\par The annual breakdown was:\\par\n")
  cat("\\begingroup\\setlength{\\parindent}{0pt}\\setlength{\\parskip}{2pt}\n")
  for (line in year_lines) {
    cat(
      "\\noindent\\hangindent=1.25em\\hangafter=1\\textbullet\\enspace ",
      line,
      "\\par\n",
      sep = ""
    )
  }
  cat("\\endgroup\n")
}

bl_pollutant_latex <- function(pollutant_md) {
  switch(
    pollutant_md,
    "NO~2~" = "NO\\textsubscript{2}",
    "PM~2.5~" = "PM\\textsubscript{2.5}",
    pollutant_md
  )
}

bl_format_marylebone_paragraph <- function(
    selection,
    pollutant_md,
    footnote_mark = NULL) {
  if (is.null(selection)) {
    return(NULL)
  }
  pol <- bl_pollutant_latex(pollutant_md)
  x <- bl_format_concentration(selection$headline_community)
  y <- bl_format_concentration(selection$headline_marylebone)

  text <- switch(
    selection$option,
    A = sprintf(
      paste0(
        "Residents may think that living outside Central London means cleaner air. ",
        "However, during the morning rush hour, the community node recorded %s µg/m³ %s, ",
        "compared with %s µg/m³ at Marylebone Road."
      ),
      x, pol, y
    ),
    B = sprintf(
      paste0(
        "Short-term %s spikes can have important health impacts. ",
        "On %d afternoons, the community node recorded a mean of %s µg/m³, ",
        "compared with %s µg/m³ at Marylebone Road."
      ),
      pol, selection$n_spike_days, x, y
    ),
    SPIKE = sprintf(
      paste0(
        "Short-term %s spikes can have important health impacts. ",
        "On %d %s, the community node recorded a mean of %s µg/m³, ",
        "compared with %s µg/m³ at Marylebone Road."
      ),
      pol, selection$n_spike_days, selection$period_label, x, y
    ),
    C = sprintf(
      paste0(
        "During school pick-up periods, the mean %s concentration at the community node was %s, ",
        "whereas the mean concentration at Marylebone Road was %s. ",
        "For families, this means that children's exposure to air pollution can be comparable ",
        "to that experienced in one of London's busiest government and traffic districts."
      ),
      pol, paste0(x, " µg/m³"), paste0(y, " µg/m³")
    ),
    NULL
  )

  if (is.null(text) || is.null(footnote_mark)) {
    return(text)
  }
  sprintf("%s%s", text, footnote_mark)
}

bl_marylebone_footnote_mark <- function(index) {
  sprintf("\\textsuperscript{%d}", index + 1L)
}

bl_collect_marylebone_footnotes <- function(entries) {
  active <- Filter(function(entry) !is.null(entry$selection), entries)
  if (!length(active)) {
    return(list(footnotes = list(), marks = list()))
  }

  footnotes <- lapply(seq_along(active), function(i) {
    entry <- active[[i]]
    list(
      mark = bl_marylebone_footnote_mark(i),
      text = {
        footnote_text <- bl_format_marylebone_footnote(
          entry$selection,
          bl_pollutant_latex(entry$pollutant_md),
          quarter_label
        )
        if (is.null(footnote_text)) {
          NULL
        } else {
          gsub("sensor", "node", footnote_text, fixed = TRUE)
        }
      }
    )
  })
  marks <- stats::setNames(
    vapply(footnotes, function(footnote) footnote$mark, character(1)),
    vapply(active, function(entry) entry$pollutant_md, character(1))
  )
  list(footnotes = footnotes, marks = as.list(marks))
}

bl_md_to_latex_inline <- function(text) {
  gsub("\\*\\*(.+?)\\*\\*", "\\\\textbf{\\1}", text, perl = TRUE)
}

bl_format_day_list <- function(day_names) {
  days <- paste0(day_names, "s")
  if (length(days) == 1L) {
    return(days[[1]])
  }
  if (length(days) == 2L) {
    return(paste(days, collapse = " and "))
  }
  paste(paste(head(days, -1L), collapse = ", "), "and", tail(days, 1L))
}

bl_format_hour_range <- function(hours) {
  hours <- sort(unique(as.integer(hours)))
  if (!length(hours)) {
    return("unknown times")
  }
  sprintf("%s to %s", bl_format_ampm_hour(min(hours)), bl_format_ampm_hour(max(hours)))
}

bl_format_ampm_hour <- function(hour) {
  hour <- as.integer(hour)
  if (hour == 0L) {
    return("12 AM")
  }
  if (hour == 12L) {
    return("12 PM")
  }
  if (hour < 12L) {
    return(sprintf("%d AM", hour))
  }
  sprintf("%d PM", hour - 12L)
}

bl_peak_hour_window <- function(day_hours, peak_frac = 0.92, max_span = 5L) {
  day_hours <- day_hours %>% arrange(hour)
  if (!nrow(day_hours)) {
    return(integer())
  }
  peak_idx <- which.max(day_hours$mean_val)
  peak <- day_hours$mean_val[peak_idx]
  threshold <- peak * peak_frac
  left <- peak_idx
  while (left > 1L && day_hours$mean_val[left - 1L] >= threshold) {
    left <- left - 1L
  }
  right <- peak_idx
  while (right < nrow(day_hours) && day_hours$mean_val[right + 1L] >= threshold) {
    right <- right + 1L
  }
  hours <- day_hours$hour[left:right]
  if (length(hours) > max_span) {
    center <- day_hours$hour[peak_idx]
    half <- floor(max_span / 2)
    hours <- seq(center - half, center + half)
    hours <- hours[hours >= 0 & hours <= 23]
  }
  hours
}

bl_prepare_peak_data <- function(df, value_col) {
  df %>%
    mutate(
      datetime = lubridate::ymd_hms(sub("\\..*", "", date)),
      day = as.Date(datetime),
      hour = lubridate::hour(datetime),
      weekday = lubridate::wday(datetime)
    ) %>%
    filter(between(day, quarter_start, quarter_end)) %>%
    filter(!is.na(.data[[value_col]]))
}

bl_day_names <- c(
  "Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday"
)

bl_summarise_record_peak <- function(period_df, value_col, weekdays_filter) {
  filtered <- period_df %>% filter(weekday %in% weekdays_filter)
  if (!nrow(filtered)) {
    return(NULL)
  }

  peak_row <- filtered %>%
    slice_max(order_by = .data[[value_col]], n = 1, with_ties = FALSE)
  peak_val <- peak_row[[value_col]][[1]]
  peak_day <- peak_row$day[[1]]
  peak_hour <- peak_row$hour[[1]]
  day_name <- bl_day_names[peak_row$weekday[[1]]]

  sprintf(
    "%s µg/m³ on %s, %d %s %s at %s",
    bl_format_concentration(peak_val),
    day_name,
    lubridate::day(peak_day),
    format(peak_day, "%B"),
    format(peak_day, "%Y"),
    bl_format_ampm_hour(peak_hour)
  )
}

bl_summarise_typical_peak <- function(period_df, value_col, weekdays_filter, period_name) {
  filtered <- period_df %>% filter(weekday %in% weekdays_filter)
  if (!nrow(filtered)) {
    return(NULL)
  }

  by_day_hour <- filtered %>%
    group_by(weekday, hour) %>%
    summarise(mean_val = mean(.data[[value_col]], na.rm = TRUE), .groups = "drop")

  day_rank <- by_day_hour %>%
    group_by(weekday) %>%
    summarise(peak_val = max(mean_val), .groups = "drop") %>%
    arrange(desc(peak_val))

  top_val <- day_rank$peak_val[[1]]
  top_days <- day_rank$weekday[day_rank$peak_val >= top_val * 0.98]
  day_label <- bl_format_day_list(bl_day_names[top_days])

  ref_hours <- by_day_hour %>%
    filter(weekday == top_days[[1]]) %>%
    arrange(hour)
  peak_hours <- bl_peak_hour_window(ref_hours)

  sprintf(
    "%s from %s have the highest concentrations during %s",
    day_label,
    bl_format_hour_range(peak_hours),
    period_name
  )
}

bl_format_record_peak_bullet <- function(period_df, value_col) {
  weekday_peak <- bl_summarise_record_peak(period_df, value_col, 2:6)
  weekend_peak <- bl_summarise_record_peak(period_df, value_col, c(1, 7))

  if (is.null(weekday_peak) && is.null(weekend_peak)) {
    return(NULL)
  }
  if (is.null(weekday_peak)) {
    return(sprintf(
      "The highest measurement during a weekend was %s.",
      weekend_peak
    ))
  }
  if (is.null(weekend_peak)) {
    return(sprintf(
      "The highest measurement during a weekday was %s.",
      weekday_peak
    ))
  }
  sprintf(
    "The highest measurement during a weekday was %s and for the weekend it was %s.",
    weekday_peak,
    weekend_peak
  )
}

bl_format_typical_peak_bullet <- function(period_df, value_col) {
  weekday_peak <- bl_summarise_typical_peak(period_df, value_col, 2:6, "weekdays")
  weekend_peak <- bl_summarise_typical_peak(period_df, value_col, c(1, 7), "weekends")

  if (is.null(weekday_peak) && is.null(weekend_peak)) {
    return(NULL)
  }
  if (is.null(weekday_peak)) {
    return(bl_append_typical_peak_footnote(sprintf("On average, %s.", weekend_peak)))
  }
  if (is.null(weekend_peak)) {
    return(bl_append_typical_peak_footnote(sprintf("On average, %s.", weekday_peak)))
  }
  bl_append_typical_peak_footnote(sprintf("On average, %s and %s.", weekday_peak, weekend_peak))
}

bl_summarise_pollutant_peaks <- function(df, value_col) {
  period_df <- bl_prepare_peak_data(df, value_col)
  list(
    record = bl_format_record_peak_bullet(period_df, value_col),
    typical = bl_format_typical_peak_bullet(period_df, value_col)
  )
}

bl_emit_summary_box <- function(
    title,
    stats,
    pollutant_md,
    threshold,
    period_label,
    border_color = "blno2quarterborder",
    peak_bullets = NULL,
    marylebone_selection = NULL,
    marylebone_footnote_mark = NULL,
    last_bullet = NULL
) {
  exceed <- bl_md_to_latex_inline(
    bl_format_who_exceedance(
      stats, bl_pollutant_latex(pollutant_md), threshold, period_label
    )
  )
  conc <- bl_format_concentration_stats(stats, period_label)

  cat("\\begin{blsummarybox}{", border_color, "}\n", sep = "")
  cat("\\textbf{", title, "}\\par\n", sep = "")
  cat("\\vspace{-0.1\\baselineskip}\n") # Adjust spacing between title and content
  cat("{\\setlength{\\topsep}{0pt}\\setlength{\\partopsep}{0pt}\\setlength{\\parsep}{0pt}\\setlength{\\itemsep}{2pt}\n")
  cat("\\begin{itemize}\n")
  cat("\\item ", exceed, "\n", sep = "")
  bl_emit_annual_exceedance_sub_bullets(stats, period_label)
  cat("\\item ", conc, "\n", sep = "")
  if (!is.null(peak_bullets)) {
    if (!is.null(peak_bullets$record)) {
      cat("\\item ", peak_bullets$record, "\n", sep = "")
    }
    if (!is.null(peak_bullets$typical)) {
      cat("\\item ", peak_bullets$typical, "\n", sep = "")
    }
  }
  marylebone_text <- bl_format_marylebone_paragraph(
    marylebone_selection, pollutant_md, marylebone_footnote_mark
  )
  if (!is.null(marylebone_text)) {
    cat("\\item ", marylebone_text, "\n", sep = "")
  }
  if (!is.null(last_bullet)) {
    cat("\\item ", last_bullet, "\n", sep = "")
  }
  cat("\\end{itemize}\n")
  cat("}\n")
  # blsummarybox ends in vertical mode, so this skip is discardable at a page
  # break instead of carrying onto the next page as a top gap.
  cat("\\end{blsummarybox}\n\\vspace{1\\baselineskip}\n\n")
}

bl_prepare_no2_pm25_pair <- function(no2_df, pm25_df, quarter_start, quarter_end) {
  no2 <- no2_df %>%
    mutate(
      datetime = lubridate::ymd_hms(sub("\\..*", "", date), quiet = TRUE),
      no2 = as.numeric(no2)
    ) %>%
    filter(!is.na(datetime), !is.na(no2)) %>%
    transmute(datetime, no2)

  pm25 <- pm25_df %>%
    mutate(
      datetime = lubridate::ymd_hms(sub("\\..*", "", date), quiet = TRUE),
      pm25 = as.numeric(pm25)
    ) %>%
    filter(!is.na(datetime), !is.na(pm25)) %>%
    transmute(datetime, pm25)

  inner_join(no2, pm25, by = "datetime") %>%
    filter(between(as.Date(datetime), quarter_start, quarter_end))
}

bl_spike_cooccurrence_metrics <- function(pair_df, quantile = 0.75, min_paired_hours = 20L) {
  # Check if the data frame is empty or does not have enough paired hours
  if (!nrow(pair_df) || nrow(pair_df) < min_paired_hours) {
    return(NULL)
  }

  # Calculate the threshold for NO2 and PM2.5 based on the specified quantile
  no2_threshold <- stats::quantile(pair_df$no2, quantile, na.rm = TRUE)
  pm25_threshold <- stats::quantile(pair_df$pm25, quantile, na.rm = TRUE)

  # Flag spikes for NO2 and PM2.5
  flagged <- pair_df %>%
    mutate(
      # Flag NO2 and PM2.5 spikes as TRUE if the concentration is greater than the threshold
      no2_spike = no2 >= no2_threshold,
      pm25_spike = pm25 >= pm25_threshold
    )

  # Count the number of NO2 spikes
  n_no2_spike <- sum(flagged$no2_spike)
  # Count the number of PM2.5 spikes
  n_pm25_spike <- sum(flagged$pm25_spike)
  # Count the number of times both spiked together
  n_both_spike <- sum(flagged$no2_spike & flagged$pm25_spike)
  # Count the number of times either spiked (union)
  n_all_spike <- sum(flagged$no2_spike | flagged$pm25_spike)

  # Return NULL if there are no spikes at all
  if (n_all_spike == 0L) {
    return(NULL)
  }

  # Return the results as a list
  list(
    n_paired_hours = nrow(pair_df),
    n_no2_spike = n_no2_spike,
    n_pm25_spike = n_pm25_spike,
    n_both_spike = n_both_spike,
    n_all_spike = n_all_spike,
    pct_both_of_all_spikes = 100 * n_both_spike / n_all_spike,
    quantile = quantile
  )
}

bl_format_spike_cooccurrence_bullet <- function(metrics, quarter_label, footnote_mark = NULL) {
  if (is.null(metrics)) {
    return(NULL)
  }

  no2_pol <- bl_pollutant_latex("NO~2~")
  pm25_pol <- bl_pollutant_latex("PM~2.5~")

  text <- sprintf(
    paste0(
      "In %s, there were %d hour%s where the highest concentration spikes in ",
      "%s and %s happened at the same time, accounting for %s of all concentration spikes measured by the community node."
    ),
    quarter_label,
    metrics$n_both_spike,
    if (metrics$n_both_spike == 1L) "" else "s",
    no2_pol,
    pm25_pol,
    bl_format_latex_pct(metrics$pct_both_of_all_spikes)
  )

  if (is.null(footnote_mark)) {
    return(text)
  }
  sprintf("%s%s", text, footnote_mark)
}

bl_spike_cooccurrence_footnote_mark <- function(marylebone_footnote_count) {
  sprintf("\\textsuperscript{%d}", marylebone_footnote_count + 2L)
}

bl_spike_cooccurrence_methodology_text <- function(quantile = BL_SPIKE_COOCCURRENCE_QUANTILE) {
  top_pct <- sprintf("%d\\%%", as.integer(100 * (1 - quantile)))
  paste(
    sprintf(
      paste0(
        "Air pollutant spikes are the highest %s of hourly concentrations for each pollutant ",
        "at this node within the report quarter, using paired measurements at the same timestamp. ",
        "The percentage reported is simultaneous spike hours as a share of all hours where either ",
        "NO\\textsubscript{2} or PM\\textsubscript{2.5} spiked."
      ),
      top_pct
    )
  )
}

if (!preview_hourly) {
  no2_quarter_stats <- bl_pollutant_summary(NO2_DATA_FRAME, "no2", WHO_NO2_DAILY, quarter_start, quarter_end)
  no2_total_stats <- bl_to_date_pollutant_summary(
    NO2_DATA_FRAME, "no2", WHO_NO2_DAILY, quarter_end, report_year, report_quarter
  )
  pm25_quarter_stats <- bl_pollutant_summary(PM25_DATA_FRAME, "pm25", WHO_PM25_DAILY, quarter_start, quarter_end)
  pm25_total_stats <- bl_to_date_pollutant_summary(
    PM25_DATA_FRAME, "pm25", WHO_PM25_DAILY, quarter_end, report_year, report_quarter
  )
  no2_peak_periods <- bl_summarise_pollutant_peaks(NO2_DATA_FRAME, "no2")
  pm25_peak_periods <- bl_summarise_pollutant_peaks(PM25_DATA_FRAME, "pm25")

  no2_pm25_pair <- bl_prepare_no2_pm25_pair(
    NO2_DATA_FRAME, PM25_DATA_FRAME, quarter_start, quarter_end
  )
  spike_cooccurrence <- bl_spike_cooccurrence_metrics(
    no2_pm25_pair, quantile = BL_SPIKE_COOCCURRENCE_QUANTILE
  )

  # Marylebone Road (MY1) reference — loads data/processed/marylebone.RData when present
  # (no openair call). If missing, fetches once, saves RData, then reuses for later reports.
  marylebone_hourly <- bl_load_marylebone_hourly()
  no2_marylebone_narrative <- bl_select_marylebone_narrative(
    NO2_DATA_FRAME, marylebone_hourly, "no2", quarter_start, quarter_end
  )
  pm25_marylebone_narrative <- bl_select_marylebone_narrative(
    PM25_DATA_FRAME, marylebone_hourly, "pm25", quarter_start, quarter_end
  )

  marylebone_footnote_entries <- list(
    list(pollutant_md = "NO~2~", selection = no2_marylebone_narrative),
    list(pollutant_md = "PM~2.5~", selection = pm25_marylebone_narrative)
  )
  marylebone_footnote_bundle <- bl_collect_marylebone_footnotes(marylebone_footnote_entries)
  marylebone_footnotes <- marylebone_footnote_bundle$footnotes
  no2_marylebone_footnote_mark <- marylebone_footnote_bundle$marks[["NO~2~"]]
  pm25_marylebone_footnote_mark <- marylebone_footnote_bundle$marks[["PM~2.5~"]]

  spike_cooccurrence_footnote_mark <- if (!is.null(spike_cooccurrence)) {
    bl_spike_cooccurrence_footnote_mark(length(marylebone_footnotes))
  } else {
    NULL
  }
  spike_cooccurrence_bullet <- bl_format_spike_cooccurrence_bullet(
    spike_cooccurrence, quarter_label, spike_cooccurrence_footnote_mark
  )
  spike_cooccurrence_methodology_text <- bl_spike_cooccurrence_methodology_text(
    BL_SPIKE_COOCCURRENCE_QUANTILE
  )

  report_date <- format(Sys.Date(), "%d %b %Y")
  lungs_banner_path <- "figure/lungs_blue.png"
  if (!file.exists(lungs_banner_path)) {
    stop("Title banner image not found: ", lungs_banner_path, call. = FALSE)
  }
}
```


```{r first-page-banner, echo=FALSE, results='asis', eval=!preview_hourly}
cat("\\thispagestyle{blfirstpage}\n")
cat(sprintf("\\renewcommand{\\blpageonetitlelineone}{%s}\n", report_title_line1))
cat(sprintf("\\renewcommand{\\blpageonetitlelinetwo}{%s}\n", report_title_line2))
cat(sprintf("\\renewcommand{\\blpageonetitlelinethree}{%s}\n", report_title_line3))
cat("\\blpageonebanner\n")
cat("\\vspace*{\\dimexpr\\blbannerreserve+\\blbannerpadding\\relax}\n")
```

```{r first-page-footer, echo=FALSE, results='asis', eval=!preview_hourly}
cat(sprintf(
  paste0(
    "\\renewcommand{\\blpageonefooter}{",
    "\\parbox{\\textwidth}{\\raggedright\\small ",
    "Breathe London Community Programme\\\\[0.3em] ",
    "Environmental Research Group, Imperial College London\\\\[0.3em] ",
    "Dr. Kayla Schulte, \\href{mailto:k.schulte21@imperial.ac.uk}{\\textcolor{blurl}{k.schulte21@imperial.ac.uk}}\\\\[0.3em] ",
    "%s",
    "}}"
  ),
  report_date
))
```

```{r org-heading, echo=FALSE, results='asis', eval=!preview_hourly}
cat(sprintf(
  paste0(
    "\\begingroup\\centering\n",
    "\\fontsize{28}{32}\\selectfont\\bfseries\\textcolor{blorg}{To the members of %s}\\par\n",
    "\\endgroup\n"
  ),
  organisation_name
))
```

```{r preview-skip-front-matter, echo=FALSE, results='asis', eval=!preview_hourly}
cat("\\vspace{2.5\\baselineskip}\n")
```

```{r sensor-location-map, echo=FALSE, fig.align='center', out.width='\\textwidth', eval=!preview_hourly}
sensor_map_path <- bl_render_site_location_map(
  site = site_code,
  icon_png = "figure/Sensor-blue-V1-removebg.png",
  icon_scale = 0.24
)
knitr::include_graphics(sensor_map_path)
```

```{r, echo=FALSE}
# Strip sub-second precision, tag day-of-week, split date/time, and format hour as HH:MM
parse_hourly <- function(df) {
  df$date    <- gsub("\\..*", "", df$date)
  df$weekday <- lubridate::wday(lubridate::ymd_hms(df$date))
  df         <- df %>% dplyr::mutate(time = sub("^\\S+\\s+", "", date))
  df         <- separate(df, date, into = c("Date2", "time2"), sep = "^\\S*\\K\\s+")
  df$Date2   <- as.Date(df$Date2)
  df$time2   <- NULL
  df <- within(df, {
    posb  <- as.POSIXlt(time, format = "%H:%M:%S")
    hours <- posb$hour
    time  <- format(posb, "%H:%M")
    posb  <- NULL
  })
  pad      <- mapply(function(x, y) paste0(rep(x, y), collapse = ""), 0, 4 - nchar(df$hours))
  df$hours <- format(strptime(paste0(pad, df$hours), format = "%M%H"), format = "%H:%M")
  df
}

# Compute diurnal means for all data and the current quarter, labelled for the plot legend
hourly_means_df <- function(df, value_col, quarter_label) {
  mean_fn <- function(x) round(mean(x, na.rm = TRUE), BL_REPORT_ROUND_DIGITS)
  means_all <- df %>% group_by(hours) %>% summarise_at(value_col, list(mean = mean_fn))
  means_quarter <- df %>%
    filter(between(Date2, quarter_start, quarter_end)) %>%
    group_by(hours) %>% summarise_at(value_col, list(mean = mean_fn))
  means_all$name     <- all_data_label
  means_quarter$name <- quarter_label
  rbind(means_all, means_quarter)
}

# Quarter means are drawn as bars, all-data means as an overlaid line
quarter_rows  <- function(df) df %>% filter(name != all_data_label)
all_data_rows <- function(df) df %>% filter(name == all_data_label)

# PM2.5 quarter bars: light grey fill + small mid-grey dots scattered in each bar
pm25_bar_dots_df <- function(df, y_step = 0.32) {
  qr <- quarter_rows(df)
  dots <- lapply(seq_len(nrow(qr)), function(i) {
    m <- qr$mean[i]
    if (!is.finite(m) || m <= 0) return(NULL)
    ys <- seq(y_step * 0.45, m - y_step * 0.25, by = y_step)
    if (!length(ys)) ys <- min(m / 2, m)
    data.frame(
      hours = rep(qr$hours[i], length(ys) * 2L),
      y = rep(ys, each = 2L),
      stringsAsFactors = FALSE
    )
  })
  dplyr::bind_rows(dots)
}

pm25_quarter_col_layer <- function(df) {
  ggplot2::geom_col(
    data = quarter_rows(df),
    aes(x = factor(hours), y = mean),
    fill = pm25_quarter_col,
    colour = pm25_quarter_dot_col,
    linewidth = 0.25,
    width = 0.78,
    position = "identity",
    inherit.aes = FALSE
  )
}

# Horizontal legend for the three overlay line series (above the hourly panels).
bl_hourly_line_legend_grob <- function(line_cols, line_types, legend_line_widths, text_size = 11) {
  dummy <- data.frame(
    x = rep(seq_len(length(line_levels)), each = 2L),
    y = rep(c(1, 1.1), length(line_levels)),
    name = factor(rep(line_levels, each = 2L), levels = line_levels)
  )
  p <- ggplot2::ggplot(
    dummy,
    ggplot2::aes(x = x, y = y, colour = name, linetype = name, linewidth = name, group = name)
  ) +
    ggplot2::geom_line() +
    ggplot2::scale_colour_manual(
      values = line_cols,
      breaks = line_levels,
      name = NULL,
      guide = ggplot2::guide_legend(
        override.aes = list(
          linetype = unname(line_types[line_levels]),
          linewidth = unname(legend_line_widths[line_levels])
        )
      )
    ) +
    ggplot2::scale_linetype_manual(values = line_types, breaks = line_levels, guide = "none") +
    ggplot2::scale_linewidth_manual(values = legend_line_widths, breaks = line_levels, guide = "none") +
    ggplot2::theme_void() +
    ggplot2::theme(
      legend.position = "top",
      legend.justification = "center",
      legend.direction = "horizontal",
      legend.text = ggplot2::element_text(size = text_size),
      legend.key.width = grid::unit(1.4, "cm"),
      legend.spacing.x = grid::unit(0.5, "cm")
    )
  cowplot::get_legend(p)
}

pm25_quarter_dot_layer <- function(df) {
  ggplot2::geom_point(
    data = pm25_bar_dots_df(df),
    aes(x = factor(hours), y = y),
    colour = pm25_quarter_dot_col,
    size = 0.22,
    position = ggplot2::position_jitter(width = 0.34, height = 0.22, seed = 42),
    inherit.aes = FALSE
  )
}

# Label y for series end text (nudge off nearby limit / other labels)
bl_label_y <- function(series_df, limit_ys = numeric(), y_max) {
  end <- series_df %>%
    dplyr::mutate(hours = as.character(hours)) %>%
    dplyr::arrange(hours) %>%
    dplyr::slice_tail(n = 1)
  y_lab <- as.numeric(end$mean) + 2.8
  gap <- 3.8
  for (pass in seq_len(5)) {
    for (ly in limit_ys) {
      if (abs(y_lab - ly) < gap) y_lab <- if (y_lab >= ly) ly + gap else ly - gap
    }
  }
  min(max(y_lab, 1.5), y_max - 0.8)
}

# Place in-panel line labels + connector sticks.
# Series order is kept as given. First letter of each label starts at start_hours
# (hjust = 0). Text and sticks stay clear of all series/bars under the label span.
bl_place_line_annots <- function(series, quarter_df, limit_ys, y_max,
                                 start_hours = c("00:00", "09:00", "16:00")) {
  stopifnot(length(series) >= 1L)
  hours <- sort(unique(unlist(lapply(series, function(s) as.character(s$df$hours)))))
  n <- length(hours)
  if (!n) stop("No hours available for line annotations", call. = FALSE)
  m <- length(series)
  if (length(start_hours) < m) {
    stop("Need at least ", m, " start_hours for line annotations", call. = FALSE)
  }

  text_width <- function(label) max(4, nchar(label) * 0.45)
  text_height <- 3.6
  line_gap <- 1.4
  label_gap <- 1.0
  stick_text_gap <- 0.45

  qd <- quarter_df %>%
    dplyr::mutate(hours = as.character(hours), bar = as.numeric(mean)) %>%
    dplyr::select(hours, bar)
  bar_y <- qd$bar[match(hours, qd$hours)]
  bar_y[!is.finite(bar_y)] <- 0

  series_y <- matrix(NA_real_, nrow = m, ncol = n)
  for (k in seq_len(m)) {
    sdf <- series[[k]]$df %>%
      dplyr::mutate(hours = as.character(hours), mean = as.numeric(mean))
    series_y[k, ] <- sdf$mean[match(hours, sdf$hours)]
  }
  all_y <- rbind(series_y, bar_y)
  env_at <- function(i0, i1) max(all_y[, i0:i1, drop = FALSE], na.rm = TRUE)

  x_overlaps <- function(i0, i1, j0, j1) i0 <= j1 && j0 <= i1

  placed <- list()
  out <- vector("list", m)

  for (k in seq_len(m)) {
    s <- series[[k]]
    i <- match(start_hours[[k]], hours)
    if (is.na(i)) {
      stop("start_hours not found in data: ", start_hours[[k]], call. = FALSE)
    }
    tw <- text_width(s$label)
    i1 <- min(n, i + max(1L, as.integer(ceiling(tw))) - 1L)
    line_y <- series_y[k, i]
    if (!is.finite(line_y)) line_y <- 0

    env_top <- env_at(i, i1)
    y_lab <- env_top + line_gap
    y_top <- y_lab + text_height

    for (ly in limit_ys) {
      if (ly >= y_lab - 0.6 && ly <= y_top + 0.6) {
        y_lab <- ly + 1.2
        y_top <- y_lab + text_height
      }
    }

    for (p in placed) {
      if (x_overlaps(i, i1, p$i, p$i1)) {
        y_lab <- max(y_lab, p$y_top + label_gap)
        y_top <- y_lab + text_height
      }
    }

    y_top <- y_lab + text_height
    stick_y0 <- line_y
    stick_y1 <- max(stick_y0 + 0.6, y_lab - stick_text_gap)

    placed[[k]] <- list(i = i, i1 = i1, y_top = y_top)
    out[[k]] <- data.frame(
      hours = hours[[i]],
      mean = y_lab,
      line_mean = line_y,
      stick_y0 = stick_y0,
      stick_y1 = stick_y1,
      label = s$label,
      colour = s$colour,
      hjust = 0,
      stringsAsFactors = FALSE
    )
  }
  out
}

# Quarter callout arrow scaled to y_max so visual offset from text matches PM2.5 layout.
bl_quarter_callout_arrow <- function(y_max, ref_y_max = y_max,
                                     text_x = 0.02, text_y = 0.35) {
  scale_y <- y_max / ref_y_max
  list(
    text_x = text_x,
    text_y = text_y,
    arrow_x = text_x + 0.05,
    arrow_y = text_y + 1.70 * scale_y,
    arrow_xend = text_x - 0.02,
    arrow_yend = text_y + 2.50 * scale_y
  )
}

# All-data diurnal means for one series (used for closest-sensor overlay)
hourly_all_means_df <- function(df, value_col, series_name) {
  mean_fn <- function(x) round(mean(x, na.rm = TRUE), BL_REPORT_ROUND_DIGITS)
  df %>%
    group_by(hours) %>%
    summarise_at(value_col, list(mean = mean_fn)) %>%
    mutate(name = series_name)
}

hour_breaks_3h <- sprintf("%02d:00", seq(0, 21, by = 3))
hour_labels_3h <- vapply(seq(0, 21, by = 3), bl_format_ampm_hour, character(1))
axis_tick_size <- 15
hourly_panel_title <- element_text(hjust = 0.5, size = rel(1.2), face = "bold")
hourly_y_title <- element_text(size = rel(1.2), face = "bold")
# Weekday and weekend data panels must stay equal width (positions 1 and 3).
hourly_panel_rel_widths <- c(1, 0.03, 1, 0.88)
no2_quarter_col <- "#C17EB9"
no2_quarter_col_alpha <- 0.75
no2_all_col <- "#A21A73"
no2_nearest_col <- "#069494"
london_col <- "#73A21A"
pm25_quarter_col <- "#D9D9D9"
pm25_quarter_dot_col <- "#888888"
pm25_all_col <- "#888888"
pm25_nearest_col <- "#069494"
pm25_london_col <- "#73A21A"

nearest_result <- bl_nearest_active_site(site_code = site_code)
nearest_site <- nearest_result$nearest
nearest_code <- as.character(nearest_site$SiteCode[1])
nearest_sensor_note <- bl_format_nearest_sensor_note(nearest_result)
nearest_label <- "Nearest node"
london_label <- "London mean"
line_levels <- c(all_data_label, nearest_label, london_label)
# Same thickness for all overlay lines
line_widths <- setNames(c(2.5, 2.5, 2.5), line_levels)
legend_line_widths <- setNames(c(3, 3, 3), line_levels)
nearest_readings <- bl_fetch_readings(site_code = nearest_code, clean = TRUE)

# Equal-weighted mean of BL_REPORT_SITE_CODES diurnal profiles (test stand-in for London)
bl_mean_across_sites_hourly <- function(parsed_list, value_col, weekdays_filter, label) {
  site_means <- lapply(parsed_list, function(df) {
    hourly_all_means_df(
      df %>% dplyr::filter(weekday %in% weekdays_filter),
      value_col,
      "tmp"
    ) %>%
      dplyr::select(hours, mean)
  })
  dplyr::bind_rows(site_means) %>%
    dplyr::group_by(hours) %>%
    dplyr::summarise(
      mean = round(mean(mean, na.rm = TRUE), BL_REPORT_ROUND_DIGITS),
      .groups = "drop"
    ) %>%
    dplyr::mutate(name = label)
}

london_site_codes <- bl_resolve_report_site_codes()
london_readings <- lapply(london_site_codes, function(code) {
  if (identical(code, site_code)) {
    list(no2 = NO2_DATA_FRAME, pm25 = PM25_DATA_FRAME)
  } else {
    bl_fetch_readings(site_code = code, clean = TRUE)
  }
})
london_no2_parsed <- lapply(london_readings, function(x) parse_hourly(x$no2))

# Parse dates and tag weekday number for both pollutants
NO2_parsed  <- parse_hourly(NO2_DATA_FRAME)
NO2_weekdays  <- NO2_parsed  %>% filter(weekday %in% 2:6)
NO2_weekends  <- NO2_parsed  %>% filter(weekday %in% c(1, 7))
NO2_plot_df            <- hourly_means_df(NO2_weekdays,  "no2",  quarter_label)
NO2_plot_weekends_df   <- hourly_means_df(NO2_weekends,  "no2",  quarter_label)

NO2_nearest_parsed <- parse_hourly(nearest_readings$no2)
NO2_nearest_weekdays_df <- hourly_all_means_df(
  NO2_nearest_parsed %>% filter(weekday %in% 2:6), "no2", nearest_label
)
NO2_nearest_weekends_df <- hourly_all_means_df(
  NO2_nearest_parsed %>% filter(weekday %in% c(1, 7)), "no2", nearest_label
)
NO2_london_weekdays_df <- bl_mean_across_sites_hourly(
  london_no2_parsed, "no2", 2:6, london_label
)
NO2_london_weekends_df <- bl_mean_across_sites_hourly(
  london_no2_parsed, "no2", c(1, 7), london_label
)
NO2_line_weekdays_df <- bind_rows(
  all_data_rows(NO2_plot_df), NO2_nearest_weekdays_df, NO2_london_weekdays_df
)
NO2_line_weekends_df <- bind_rows(
  all_data_rows(NO2_plot_weekends_df), NO2_nearest_weekends_df, NO2_london_weekends_df
)
NO2_line_cols <- setNames(scales::alpha(c(no2_all_col, no2_nearest_col, london_col), 0.65), line_levels)
NO2_line_types <- setNames(c("solid", "solid", "solid"), line_levels)

PM25_parsed <- parse_hourly(PM25_DATA_FRAME)
PM25_weekdays <- PM25_parsed %>% filter(weekday %in% 2:6)
PM25_weekends <- PM25_parsed %>% filter(weekday %in% c(1, 7))
PM2.5_plot_df          <- hourly_means_df(PM25_weekdays, "pm25", quarter_label)
PM2.5_plot_weekends_df <- hourly_means_df(PM25_weekends, "pm25", quarter_label)

PM25_nearest_parsed <- parse_hourly(nearest_readings$pm25)
PM25_nearest_weekdays_df <- hourly_all_means_df(
  PM25_nearest_parsed %>% filter(weekday %in% 2:6), "pm25", nearest_label
)
PM25_nearest_weekends_df <- hourly_all_means_df(
  PM25_nearest_parsed %>% filter(weekday %in% c(1, 7)), "pm25", nearest_label
)
london_pm25_parsed <- lapply(london_readings, function(x) parse_hourly(x$pm25))
PM25_london_weekdays_df <- bl_mean_across_sites_hourly(
  london_pm25_parsed, "pm25", 2:6, london_label
)
PM25_london_weekends_df <- bl_mean_across_sites_hourly(
  london_pm25_parsed, "pm25", c(1, 7), london_label
)
PM25_line_weekdays_df <- bind_rows(
  all_data_rows(PM2.5_plot_df), PM25_nearest_weekdays_df, PM25_london_weekdays_df
)
PM25_line_weekends_df <- bind_rows(
  all_data_rows(PM2.5_plot_weekends_df), PM25_nearest_weekends_df, PM25_london_weekends_df
)
PM25_line_cols <- setNames(scales::alpha(c(pm25_all_col, pm25_nearest_col, pm25_london_col), 0.65), line_levels)
PM25_line_types <- setNames(c("solid", "solid", "solid"), line_levels)

hourly_callout_ref_y_max <- max(
  25,
  max(PM2.5_plot_df$mean, PM25_line_weekdays_df$mean, PM25_line_weekends_df$mean, na.rm = TRUE),
  na.rm = TRUE
)

```

```{r no2-page-heading, echo=FALSE, results='asis', eval=!preview_hourly}
cat("\\newpage\n\n# NO~2~\n\n## Average NO~2~ for each hour at your Breathe London Community node\n")
```

```{r preview-no2-heading, echo=FALSE, results='asis', eval=preview_no2}
if (preview_no2) {
  cat("\\thispagestyle{plain}\n")
} else {
  cat("\\newpage\n")
}
cat("# NO~2~\n\n## Average NO~2~ for each hour at your Breathe London Community node\n")
```

```{r no2-hourly-plot, eval=!preview_pm25, out.width='\\textwidth', fig.width = 12.5, fig.height = 4.4, fig.align='center', fig.show='hold', fig.pos='H'}
NO2_y_max <- max(
  40,
  max(NO2_plot_df$mean, NO2_line_weekdays_df$mean, NO2_line_weekends_df$mean, na.rm = TRUE),
  na.rm = TRUE
)
NO2_line_legend <- bl_hourly_line_legend_grob(NO2_line_cols, NO2_line_types, legend_line_widths)
annot_size <- 4.9
axis_x_theme <- theme(
  axis.text.x = element_text(
    angle = 45, vjust = 1, hjust = 1, size = axis_tick_size,
    margin = margin(t = 0)
  ),
  axis.ticks.length.x = unit(0, "pt"),
  axis.title.x = element_text(margin = margin(t = 2))
)
NO2_plot <- ggplot(NO2_plot_df, aes(factor(hours), mean))+
  geom_col(data = quarter_rows(NO2_plot_df), aes(fill=name), position="identity")+
  geom_line(
    data = NO2_line_weekdays_df,
    aes(colour = name, linetype = name, linewidth = name, group = name)
  )+
  theme_minimal() +
  axis_x_theme +
  theme(
    axis.text.y = element_text(size = axis_tick_size),
    axis.title.y = hourly_y_title,
    legend.position = "none",
    plot.title = hourly_panel_title,
    plot.margin = margin(5, 1, 8, 5)
  ) +
  xlab("Hour (Weekday)") + ylab(expression(Mean~NO[2]~"in ug/m"^3)) + ggtitle("WEEKDAYS") +
  geom_hline(yintercept = 40, linetype = "dashed", color = "black") +
  geom_hline(yintercept = 25, linetype = "dashed", color = "black") +
  geom_hline(yintercept = 10, linetype = "dashed", color = "black") +
  coord_cartesian(ylim = c(0, NO2_y_max)) +
  scale_x_discrete(breaks = hour_breaks_3h, labels = hour_labels_3h) +
  scale_fill_manual(
    values = setNames(scales::alpha(no2_quarter_col, no2_quarter_col_alpha), quarter_label),
    name = "Data range"
  ) +
  scale_colour_manual(values = NO2_line_cols, breaks = line_levels, name = NULL) +
  scale_linetype_manual(values = NO2_line_types, breaks = line_levels, guide = "none") +
  scale_linewidth_manual(values = line_widths, breaks = line_levels, guide = "none")

NO2_plot_weekend <- ggplot(NO2_plot_weekends_df, aes(factor(hours), mean))+
  geom_col(data = quarter_rows(NO2_plot_weekends_df), aes(fill=name), position="identity")+
  geom_line(
    data = NO2_line_weekends_df,
    aes(colour = name, linetype = name, linewidth = name, group = name)
  )+
  geom_hline(yintercept = 40, linetype = "dashed", color = "black") +
  geom_hline(yintercept = 25, linetype = "dashed", color = "black") +
  geom_hline(yintercept = 10, linetype = "dashed", color = "black") +
  theme_minimal() +
  axis_x_theme +
  theme(
    axis.title.y = element_blank(),
    axis.text.y = element_text(color = NA, size = axis_tick_size),
    axis.ticks.y = element_blank(),
    legend.position = "none",
    plot.title = hourly_panel_title,
    plot.margin = margin(5, 0, 8, 0)
  ) +
  xlab("Hour (Weekend)") + ylab(NULL) + ggtitle("WEEKENDS") +
  coord_cartesian(ylim = c(0, NO2_y_max)) +
  scale_x_discrete(breaks = hour_breaks_3h, labels = hour_labels_3h) +
  scale_fill_manual(
    values = setNames(scales::alpha(no2_quarter_col, no2_quarter_col_alpha), quarter_label),
    name = "Data range"
  ) +
  scale_colour_manual(values = NO2_line_cols, breaks = line_levels, name = NULL) +
  scale_linetype_manual(values = NO2_line_types, breaks = line_levels, guide = "none") +
  scale_linewidth_manual(values = line_widths, breaks = line_levels, guide = "none")

NO2_callout <- bl_quarter_callout_arrow(NO2_y_max, hourly_callout_ref_y_max)
NO2_limit_labels <- ggplot() +
  annotate("text", x = 0, y = 40, label = expression(paste("UK ", NO[2], " annual mean limit (40 ug/m"^3, ")")), hjust = 0, size = annot_size) +
  annotate("text", x = 0, y = 25, label = expression(paste("WHO ", NO[2], " daily mean limit (25 ug/m"^3, ")")), hjust = 0, size = annot_size) +
  annotate("text", x = 0, y = 10, label = expression(paste("WHO ", NO[2], " annual mean limit (10 ug/m"^3, ")")), hjust = 0, size = annot_size) +
  annotate(
    "text", x = NO2_callout$text_x, y = NO2_callout$text_y,
    label = expression(paste("Here is what ", NO[2], " looked like last quarter")),
    hjust = 0, vjust = 0, size = annot_size, colour = no2_quarter_col
  ) +
  annotate(
    "curve",
    x = NO2_callout$arrow_x, y = NO2_callout$arrow_y,
    xend = NO2_callout$arrow_xend, yend = NO2_callout$arrow_yend,
    curvature = 0.45,
    colour = no2_quarter_col,
    linewidth = 0.65,
    lineend = "round",
    arrow = arrow(length = unit(0.16, "cm"), type = "closed", angle = 25)
  ) +
  coord_cartesian(ylim = c(0, NO2_y_max), xlim = c(0, 1), clip = "off") +
  theme_void() +
  theme(plot.margin = margin(22, 2, 36, -10))

plot_grid(
  plot_grid(NULL, NO2_line_legend, NULL, ncol = 3, rel_widths = c(1, 2, 1)),
  NULL,
  plot_grid(
    NO2_plot, NULL, NO2_plot_weekend, NO2_limit_labels,
    align = "h", axis = "tb", ncol = 4,
    rel_widths = hourly_panel_rel_widths
  ),
  ncol = 1,
  rel_heights = c(0.07, 0.07, 1)
)
if (preview_no2) knitr::knit_exit()
```

```{r no2-nearest-sensor-note, echo=FALSE, results='asis', eval=!preview_hourly}
bl_emit_nearest_sensor_note(nearest_sensor_note)
```

```{r no2-quarter-summary, echo=FALSE, results='asis', eval=!preview_hourly}
bl_emit_summary_box(
  sprintf("Summary of %s", quarter_label),
  no2_quarter_stats,
  "NO~2~",
  WHO_NO2_DAILY,
  quarter_label,
  border_color = "blno2quarterborder",
  peak_bullets = no2_peak_periods,
  marylebone_selection = no2_marylebone_narrative,
  marylebone_footnote_mark = no2_marylebone_footnote_mark,
  last_bullet = spike_cooccurrence_bullet
)
```

```{r no2-total-summary, echo=FALSE, results='asis', eval=!preview_hourly}
bl_emit_summary_box(
  summary_to_date_heading,
  no2_total_stats,
  "NO~2~",
  WHO_NO2_DAILY,
  "To date",
  border_color = "blno2totalborder"
)
```

```{r preview-pm25-heading, echo=FALSE, results='asis', eval=preview_pm25}
if (preview_pm25) {
  cat("\\thispagestyle{plain}\n")
} else {
  cat("\\newpage\n")
}
cat("# PM~2.5~\n\n## Average PM~2.5~ for each hour at your Breathe London Community node\n")
```

```{r pm25-page-heading, echo=FALSE, results='asis', eval=!preview_pm25 && !preview_no2}
cat("# PM~2.5~\n\n## Average PM~2.5~ for each hour at your Breathe London Community node\n")
```

```{r pm25-hourly-plot, eval=!preview_no2, out.width='\\textwidth', fig.width = 12.5, fig.height = 4.4, fig.align='center', fig.show='hold', fig.pos='H'}
PM25_y_max <- max(
  25,
  max(PM2.5_plot_df$mean, PM25_line_weekdays_df$mean, PM25_line_weekends_df$mean, na.rm = TRUE),
  na.rm = TRUE
)
PM25_line_legend <- bl_hourly_line_legend_grob(PM25_line_cols, PM25_line_types, legend_line_widths)
annot_size <- 4.9
axis_x_theme <- theme(
  axis.text.x = element_text(
    angle = 45, vjust = 1, hjust = 1, size = axis_tick_size,
    margin = margin(t = 0)
  ),
  axis.ticks.length.x = unit(0, "pt"),
  axis.title.x = element_text(margin = margin(t = 2))
)
PM2.5_plot_weekdays <- ggplot(PM2.5_plot_df, aes(factor(hours), mean))+
  pm25_quarter_col_layer(PM2.5_plot_df)+
  pm25_quarter_dot_layer(PM2.5_plot_df)+
  geom_line(
    data = PM25_line_weekdays_df,
    aes(colour = name, linetype = name, linewidth = name, group = name)
  )+
  theme_minimal() +
  axis_x_theme +
  theme(
    axis.text.y = element_text(size = axis_tick_size),
    axis.title.y = hourly_y_title,
    legend.position = "none",
    plot.title = hourly_panel_title,
    plot.margin = margin(5, 1, 8, 5)
  ) +
  xlab("Hour (Weekday)") + ylab(expression(Mean~PM[2.5]~"in ug/m"^3)) + ggtitle("WEEKDAYS") +
  geom_hline(yintercept = 25, linetype = "dashed", color = "black") +
  geom_hline(yintercept = 15, linetype = "dashed", color = "black") +
  geom_hline(yintercept = 5, linetype = "dashed", color = "black") +
  coord_cartesian(ylim = c(0, PM25_y_max)) +
  scale_x_discrete(breaks = hour_breaks_3h, labels = hour_labels_3h) +
  scale_colour_manual(values = PM25_line_cols, breaks = line_levels, name = NULL) +
  scale_linetype_manual(values = PM25_line_types, breaks = line_levels, guide = "none") +
  scale_linewidth_manual(values = line_widths, breaks = line_levels, guide = "none")

PM2.5_plot_weekends <- ggplot(PM2.5_plot_weekends_df, aes(factor(hours), mean))+
  pm25_quarter_col_layer(PM2.5_plot_weekends_df)+
  pm25_quarter_dot_layer(PM2.5_plot_weekends_df)+
  geom_line(
    data = PM25_line_weekends_df,
    aes(colour = name, linetype = name, linewidth = name, group = name)
  )+
  geom_hline(yintercept = 25, linetype = "dashed", color = "black") +
  geom_hline(yintercept = 15, linetype = "dashed", color = "black") +
  geom_hline(yintercept = 5, linetype = "dashed", color = "black") +
  theme_minimal() +
  axis_x_theme +
  theme(
    axis.title.y = element_blank(),
    axis.text.y = element_text(color = NA, size = axis_tick_size),
    axis.ticks.y = element_blank(),
    legend.position = "none",
    plot.title = hourly_panel_title,
    plot.margin = margin(5, 0, 8, 0)
  ) +
  xlab("Hour (Weekend)") + ylab(NULL) + ggtitle("WEEKENDS") +
  coord_cartesian(ylim = c(0, PM25_y_max)) +
  scale_x_discrete(breaks = hour_breaks_3h, labels = hour_labels_3h) +
  scale_colour_manual(values = PM25_line_cols, breaks = line_levels, name = NULL) +
  scale_linetype_manual(values = PM25_line_types, breaks = line_levels, guide = "none") +
  scale_linewidth_manual(values = line_widths, breaks = line_levels, guide = "none")

PM25_callout <- bl_quarter_callout_arrow(PM25_y_max, hourly_callout_ref_y_max)
PM2.5_limit_labels <- ggplot() +
  annotate("text", x = 0, y = 25, label = expression(paste("UK ", PM[2.5], " annual mean limit (25 ug/m"^3, ")")), hjust = 0, size = annot_size) +
  annotate("text", x = 0, y = 15, label = expression(paste("WHO ", PM[2.5], " daily mean limit (15 ug/m"^3, ")")), hjust = 0, size = annot_size) +
  annotate("text", x = 0, y = 5, label = expression(paste("WHO ", PM[2.5], " annual mean limit (5 ug/m"^3, ")")), hjust = 0, size = annot_size) +
  annotate(
    "text", x = PM25_callout$text_x, y = PM25_callout$text_y,
    label = expression(paste("Here is what ", PM[2.5], " looked like last quarter")),
    hjust = 0, vjust = 0, size = annot_size, colour = pm25_quarter_dot_col
  ) +
  annotate(
    "curve",
    x = PM25_callout$arrow_x, y = PM25_callout$arrow_y,
    xend = PM25_callout$arrow_xend, yend = PM25_callout$arrow_yend,
    curvature = 0.45,
    colour = pm25_quarter_dot_col,
    linewidth = 0.65,
    lineend = "round",
    arrow = arrow(length = unit(0.16, "cm"), type = "closed", angle = 25)
  ) +
  coord_cartesian(ylim = c(0, PM25_y_max), xlim = c(0, 1), clip = "off") +
  theme_void() +
  theme(plot.margin = margin(22, 2, 36, -10))

plot_grid(
  plot_grid(NULL, PM25_line_legend, NULL, ncol = 3, rel_widths = c(1, 2, 1)),
  NULL,
  plot_grid(
    PM2.5_plot_weekdays, NULL, PM2.5_plot_weekends, PM2.5_limit_labels,
    align = "h", axis = "tb", ncol = 4,
    rel_widths = hourly_panel_rel_widths
  ),
  ncol = 1,
  rel_heights = c(0.07, 0.07, 1)
)
if (preview_pm25) knitr::knit_exit()
```

```{r pm25-nearest-sensor-note, echo=FALSE, results='asis', eval=!preview_hourly}
bl_emit_nearest_sensor_note(nearest_sensor_note)
```

```{r pm25-quarter-summary, echo=FALSE, results='asis', eval=!preview_hourly}
bl_emit_summary_box(
  sprintf("Summary of %s", quarter_label),
  pm25_quarter_stats,
  "PM~2.5~",
  WHO_PM25_DAILY,
  quarter_label,
  border_color = "blpm25quarterborder",
  peak_bullets = pm25_peak_periods,
  marylebone_selection = pm25_marylebone_narrative,
  marylebone_footnote_mark = pm25_marylebone_footnote_mark
)
```

```{r pm25-total-summary, echo=FALSE, results='asis', eval=!preview_hourly}
bl_emit_summary_box(
  summary_to_date_heading,
  pm25_total_stats,
  "PM~2.5~",
  WHO_PM25_DAILY,
  "To date",
  border_color = "blpm25totalborder"
)
```

\newpage

# Annual WHO Exceedance

Each year the community node measured pollution is represented by a pair of lungs. The left lung represents NO~2~ and the right PM~2.5~, with the fill level indicating the percentage of days that exceeded the WHO daily guideline. The number above each lung is the count of exceedance days that year.

```{r lung-plot-size, echo=FALSE}
# fig.height must be known before the graphics device opens, so the row count is
# resolved here and referenced from the next chunk's header.
LUNG_YEARS_PER_ROW <- 3L
lung_plot_df <- bl_lung_exceedance_year_df(no2_total_stats, pm25_total_stats)
lung_n_years <- if (is.null(lung_plot_df)) 0L else nrow(lung_plot_df)
lung_fig_height <- bl_lung_plot_fig_height(
  max(lung_n_years, 1L),
  years_per_row = LUNG_YEARS_PER_ROW
)
```

```{r lung-annual-exceedance, out.width='\\textwidth', fig.width = 11, fig.height = lung_fig_height, fig.align='center', fig.pos='H', echo=FALSE}
# Lungs-report label sizes (defaults unchanged in bl_plot_lung_exceedance() elsewhere).
lungs_text_sizes <- bl_lung_plot_text_sizes(
  year = 14,
  pct = 8,
  days = 5,
  pollutant = 6
)

lung_exceedance_plot <- bl_plot_lung_exceedance(
  no2_total_stats,
  pm25_total_stats,
  no2_color = no2_all_col,
  pm25_color = pm25_all_col,
  text_sizes = lungs_text_sizes,
  years_per_row = LUNG_YEARS_PER_ROW
)
if (!is.null(lung_exceedance_plot)) {
  print(lung_exceedance_plot)
}
```

**What is air pollution and where does it come from?**

Air Pollution includes any gases or particles in the air that can harm our health. In London, we focus on monitoring three main types of pollution:

Nitrogen dioxide (NO~2~) - This is a gas produced by burning fossil fuels. In London, the biggest source of NO~2~ is road transport at around 45%, followed by industrial and commercial heating at around 20% ([Source: LAEI, 2023](https://data.london.gov.uk/download/e758q/06aab8a6-79a6-40ae-8038-8303ac82a3aa/LAEI%202019%20Summary%20Note.pdf)). Exposure to NO~2~ has been associated with a range of health effects including organ development during pregnancy, new onset asthma in children and lung function decline in older adults.

Fine particles (PM~2.5~) - Fine particles, also known as PM~2.5~ (Particulate Matter 2.5 microns in diameter) come from a wide range of sources. In London, the greatest contributors include road transport (30%) followed by domestic wood burning at 17% and commercial cooking at 15% ([Source: LAEI, 2023](https://data.london.gov.uk/download/e758q/06aab8a6-79a6-40ae-8038-8303ac82a3aa/LAEI%202019%20Summary%20Note.pdf)). Their small size means they can enter the bloodstream and have serious health effects including increasing risk of lung cancer, heart disease, asthma, and low birth weight.

```{r report-footer, echo=FALSE, results='asis'}
cat("\\newpage\n\\thispagestyle{plain}\n\\raggedright\n")
cat("\\noindent{\\large\\bfseries General Air Pollution Trends in London\\par}\n")
cat("\\vspace{0.5\\baselineskip}\n")
for (q in seq_along(quarter_trend_texts)) {
  cat(bl_format_quarter_line(q, quarter_trend_texts[q]), "\n\n")
}
cat("\\vfill\n")
cat("\\noindent{\\fontsize{0.8em}{0.8em}\\selectfont\n")
cat("\\textsuperscript{1} ", bl_typical_peak_methodology_text, "\n\n", sep = "")
for (footnote in marylebone_footnotes) {
  cat(footnote$mark, " ", footnote$text, "\n\n", sep = "")
}
if (!is.null(spike_cooccurrence_footnote_mark)) {
  cat(spike_cooccurrence_footnote_mark, " ", spike_cooccurrence_methodology_text, "\n\n", sep = "")
}
cat(
  "To download the data that was used to generate this report, visit ",
  "\\url{https://www.breathelondon-communities.org/} and search for your community node in the ",
  "Node Search hyperlink just below the node map.\n",
  sep = ""
)
cat("}\n")
```

